### 2014-05-12 research/paper/main/article.tex
revision file added

@@

The base quasi-steady state (QSS) model was first developed by \citet{Parkinson1964} for a square cross section. The equation of motion of the body is given by 
\begin{equation}
\label{equationofmotion}
m\ddot{y}+c\dot{y}+ky=F_y,
\end{equation}
where the forcing term $F_y$ is given by
\begin{equation}
\label{lift equation}
F_y=\frac{1}{2}\rho U^2\mathcal{A}C_y.
\end{equation}
\citet{lighthill1986} showed that for systems oscillating in fluid, it is sometimes useful to decompose the fluid forces into components that are in and out of phase with the body acceleration. The component in phase with the acceleration effectively adds to the inertia or effective mass of the system. Therefore, an added mass term, $m_a$, can be added to the system mass. For consistency with previous studies such as \citet{Joly2012}, a value of $m_a=3.5$ has been used here.

@@

In the QSS model, it is assumed that the force on the body at a given instantaneous incident angle $\theta$ (defined in figure \ref{fig:setup_1}) is the same as the mean force on a static body at the same incident angle, or angle of attack. The instantaneous value of $C_y$ is therefore determined by an interpolating polynomial based on the lift data for flow over a stationary body at various $\theta$. Using the relationship between $\theta$ and the instantaneous transverse velocity of the body $\dot{y}$ shown in figure \ref{fig:setup_1}, $C_y$ can be written as a function of $\dot{y}$. The order of the interpolation polynomial used to define this function has varied from study to study. For  example a $7^{th}$ order polynomial was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order doesn't provides a significantly better result. Thus a $7 ^{th}$ order interpolating polynomial is used in this present study. As a result, $C_y(\dot{y})$ is defined as
\begin{equation}
\label{cy ploynomial}
C_y(\theta)=a_1\left(\frac{\dot{y}}{U}\right)+a_3\left(\frac{\dot{y}}{U}\right)^3+a_5\left(\frac{\dot{y}}{U}\right)^5+a_7\left(\frac{\dot{y}}{U}\right)^7.
\end{equation}

>>>>

\section{Introduction}

Fluid-elastic galloping is one of the sub-areas of research in fluid structure interactions. This area has been of interest due to the vibrations crated by galloping on transmission lines and civil structures and leading them to failure. Therefore understanding this phenomenon in order to suppresses these vibrations was quite important. However, the search for alternate energy sources with minimal environmental impact has become an important area of research in the modern word. Therefore researchers are moving towards investigating the possibility of extracting useful energy from this vibrations rather than suppressing them. Therefore, it is quite important to understand the governing parameters and analyse the influence of them on the energy transfer from the fluid to the body, because this understanding will lead to develop better practical applications. Therefore, in this paper we focus on understanding the energy transfer from the fluid to the body and isolate the governing parameters influencing it.

According to \citet{Paidoussis2010}, \citet{Glauert1919} provided a criterion for galloping by considering the auto-rotation of an aerofoil.  \citet{DenHartog1956} provided a theoretical explanation for galloping for iced electric transmission lines. A weakly non-linear theoretical aeroelastic model to predict the response of galloping was developed by \citet{Parkinson1964} based on the quasi-steady state hypothesis. Experimental lift and drag data on a fixed square prism at different angles of attack were used as an input for the theoretical model. It essentially used a curve fit of the transverse force to predict the galloping response. The study managed to achieve a good agreement with experimental data.

However, the QSS model equation when solved analytically using the sinusoidal solution method cannot predict the response for cases with low mass ratios. \citet{Joly2012} observed that finite element simulations show a sudden change in amplitude below a critical value of the mass ratio. The model equation defined in \citet{Parkinson1964} was modified to account for the vortex shedding and solved numerically to predict the reduced amplitude at low mass ratios to the point where galloping is no longer present. \citet{Barrero-Gil2010a} investigated the possibility of extracting power from vibrations caused by galloping using the quasi-steady state model. In the conclusions of that paper it was pointed out that in order to obtain a high power to area ratio, the mass-damping ($m^*\zeta$) parameter should be kept low. The same study investigated the influence of the characteristics of the $C_y$ curve on maximum power output.

Here, the modified QSS model developed by \citet{Joly2012} is integrated numerically for low Reynolds numbers. The focus is on the power extraction potential as a function of mechanical parameters (i.e. frequency of oscillation, damping factor and mass ratio). To this end, a series of previously mentioned mechanical parameters are tested at two different values of \reynoldsnumber: $\reynoldsnumber = 165$, a case that should remain laminar and essentially two-dimensional; $\reynoldsnumber = 22300$, a case where the flow is expected to be turbulent and three-dimensional. Both cases require the input of transverse force coefficients $C_y$ as a function of angle of attack $\theta$ for a fixed body. These data are provided from direct numerical simulations for the $\reynoldsnumber = 165$ case, while the data provided by \citet{Parkinson1964} are used for the $\reynoldsnumber = 22300$ case.

The structure of the paper is as follows. Section \ref{sec:theory} presents the modified QSS model, the method for the calculation of power output, and the parameters used. Section \ref{sec:results} presents the results, first of the fixed body tests at a range of $\theta$, then of the response characteristics predicted by the integration of the QSS model for both the high and low \reynoldsnumber\ cases. For the low \reynoldsnumber\ case, the results of the QSS model are compared to those of full direct numerical simulations of the fluid-structure interaction problem. Finally, section \ref{sec:conc} presents the conclusions that can be drawn from this work.
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section*{Nomenclature}
%\textbf{Nomenclature}

\begin{tabular}{ll}
$a_1,a_3,a_5,a_7$ & coefficients of the polynomial to determine $C_y$ \\ 
$A$ & displacement amplitude\\
$c$ & damping constant \\
$D$ & characteristic length (side length) of the cross section of the body \\
$f=\sqrt{k/m}/2\pi$ & natural frequency of the system \\
$F_y$ & instantaneous force normal to the flow \\ 
$F_0$& amplitude of the oscillatory force due to vortex shedding \\
$k$ & spring constant \\
$m$ & mass of the body \\
$m_a$ & added mass \\
$P_d$ & power dissipated due to mechanical damping  \\
$P_{in}=\rho U^3D/2$ & Energy flux of the approaching flow \\
$P_{mean}$ & mean power \\
$P_t$   & power transferred to the body by the fluid \\
$t$ & time \\
$U$ & freestream velocity \\
$U_i$ & Induced velocity \\
$y,\dot{y},\ddot{y}$ & transverse displacement, velocity and acceleration of the body \\
$\mathcal{A}=DL$ & frontal area of the body\\ 
$\lambda$ & Inverse time scale of a galloping dominated flow \\
$\lambda_{1,2}$ & Eigenvalues of linearized equation of motion \\
$\rho$ & fluid density  \\
$\omega_n= 2 \pi f$ & natural angular frequency of the system  \\
$\omega_s$ & vortex shedding angular frequency \\
$\cstar=cD/mU$ & non-dimensionalised damping factor \\
$C_y=F_y/0.5\rho U^2DL$ & normal (lift) force coefficient \\
$m^*=m/\rho D^2L$ & mass ratio \\
$Re$ & Reynolds number  \\
$U^*=U/fD$ & reduced velocity  \\
$Y=y/D$ & non-dimensional transverse displacement \\
$\dot{Y}=m^*\dot{y}/a_1U$ & non-dimensional transverse velocity \\
$\ddot{Y}=m^{*2}D/a_1^2U^2$ & non-dimensional transverse acceleration \\
$\Gamma_1 = 4\pi^2m^{*2}/U^{*2}a_1^2$ & First dimensionless group arising from linearised, non-dimensionalised equation of motion\\
$\Gamma_2 = c^*m^*/a_1$ & Second dimensionless group arising from linearised, non-dimensionalised equation of motion\\
$\zeta= c/2 m \omega_n$ & damping ratio \\
$\theta= \tan^{-1}{(\dot{y}/U)}$ & instantaneous angle of incidence (angle of attack)\\
$\massstiff =  4\pi^2m^{*2}/U^{*2}$ & Combined mass-stiffness parameter\\
$\massdamp = c^*m^*$ & Combined mass-damping parameter\\
\end{tabular}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

@@

The equation of motion of the body is given by 
\begin{equation}
\label{equationofmotion}
(m+m_a)\ddot{y}+c\dot{y}+ky=F_y,
\end{equation}
where the forcing term $F_y$ is given by
\begin{equation}
\label{lift equation}
F_y=\frac{1}{2}\rho U^2\mathcal{A}C_y.
\end{equation}
\citet{lighthill1986} showed that for systems oscillating in fluid, it is sometimes useful to decompose the fluid forces into components that are in and out of phase with the body acceleration. The component in phase with the acceleration effectively adds to the inertia or effective mass of the system. Therefore, an added mass term, $m_a$, can be added to the system mass. For consistency with previous studies such as \citet{Joly2012}, a value of $m_a=3.5$ has been used here.

@@

In the QSS model, it is assumed that the force on the body at a given instantaneous incident angle $\theta$ (defined in figure \ref{fig:setup_1}) is the same as the mean force on a static body at the same incident angle, or angle of attack. The instantaneous value of $C_y$ is therefore determined by an interpolating polynomial based on the lift data for flow over a stationary body at various $\theta$. Using the relationship between $\theta$ and the instantaneous transverse velocity of the body $\dot{y}$ shown in figure \ref{fig:setup_1}, $C_y$ can be written as a function of $\dot{y}$. The order of the interpolation polynomial used to define this function has varied from study to study. For  example a $7^{th}$ order polynomial was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order doesn't provides a significantly better result. Thus a $7 ^{th}$ order interpolating polynomial is used in this present study. As a result, $C_y(\theta)$ (noting that theta is proportional to $\dot{y}/U$) is defined as
\begin{equation}
\label{cy ploynomial}
C_y(\theta)=a_1\left(\frac{\dot{y}}{U}\right)+a_3\left(\frac{\dot{y}}{U}\right)^3+a_5\left(\frac{\dot{y}}{U}\right)^5+a_7\left(\frac{\dot{y}}{U}\right)^7.
\end{equation}
